### 2021-07-06 paper1/paper1.tex
Address the remaining feedback
Fix all minor issues

There are many existing models for enforcing demographic parity and equalised odds
\citep{creager2019flexibly,AgaBeyDudLanetal18,calders2009building,kamishima2012fairness,zafar2017fairnesstreatment,zafar2017fairnessconstraints}.
However, these existing approaches to balancing accuracy and fairness rely on intermediate, unintuitive control parameters such as allowable constraint violation $\epsilon$ (e.g. $0.01$) in \citet{AgaBeyDudLanetal18}, or a covariance threshold $c$
(e.g. $0$ that is controlled by another parameters $\tau$ and $\mu$ -- $0.005$ and $1.2$ -- to trade off this threshold and accuracy)
in \citet{zafar2017fairnesstreatment}.
This is related to the fact that many of these approaches embed fairness criteria as \emph{constraints} in the optimisation procedure
\citep{DonOneBenShaetal18,quadrianto2017recycling,zafar2017fairnesstreatment,zafar2017fairnessconstraints}.

@@

\section{Target labels for tuning group fairness}
We will start by describing several notions of group fairness.
For each individual, we have a vector of non-sensitive attributes $x\in\mathcal{X}$, a class label $y\in\mathcal{Y}$, and a sensitive attribute $s\in\mathcal{S}$ (e.g.\ racial origin or gender).
%
We focus on the case where $s$ and $y$ are binary. 
%
We assume that a positive label $y=1$ corresponds to a positive outcome for an individual -- for example, being accepted for a loan.
%
%
\emph{Group fairness} balances a certain condition between groups of individuals with different sensitive attribute, $s$ versus $s\sp{\prime}$. 
%
The term $\hat{y}$ below is the prediction of a machine learning model that,
in most works, uses only non-sensitive attributes $x$.
%
Several group fairness criteria have been proposed (e.g. \cite{zafar2017fairnesstreatment,chouldechova2017fair,hardt2016equality}):
\begin{align}
& \text{equality of positive rate (Demographic Parity):}\nonumber\\
& P(\hat{y}=1|s)=P(\hat{y}=1|s\sp{\prime}) \label{eq:eq_dem_par}\\
& \text{equality of accuracy:}\nonumber\\
& P(\hat{y}=y|s)=P(\hat{y}=y|s\sp{\prime}) \\
& \text{equality of true positive rate (Equality of Opportunity):} \nonumber\\
& P(\hat{y}=1|s,y=1)=P(\hat{y}=1|s\sp{\prime},y=1)  \label{eq:eq_opp_cri}~.
\end{align}
\emph{Equalised odds} criterion corresponds to Equality of Opportunity (\ref{eq:eq_opp_cri}) plus equality of false positive rate.

@@

%We now present a novel method for learning from an only implicitly-defined balanced dataset.
We present a novel approach which only implicitly constructs the balanced dataset.
This framework can be used with any likelihood-based model, such as Logistic Regression and Gaussian Process models.
The relation presented in \eqref{eq:etaeta} allows us to formulate a likelihood
that targets $\bar{\eta}_s(x)$ while only having access to the imbalanced labels $y$.
As we only have access to $y$, $P(y|x,s,\theta)$ is the likelihood to optimise.
It represents the probability that $y$ is the imbalanced label,
given the input $x$, the sensitive attribute $s$ that available in the training set
and the model parameters $\theta$ for a model that is targeting $\bar{y}$.
Thus, we get
\begin{align}
  &P(y=1|x, s, \theta)
  = \sum\limits_{\bar{y}\in \{0, 1\}} P(y=1,\bar{y}|x, s, \theta) \nonumber\\
  =\,\, &\sum\limits_{\bar{y}\in \{0, 1\}} P(y=1|\bar{y}, x, s, \theta)\,P(\bar{y}|x, s, \theta) \label{eq:lik}~.
\end{align}
As we are only considering group fairness, we have $P(y=1|\bar{y}, x, s, \theta)=P(y=1|\bar{y}, s)$.

@@

The previous statements about $t_s$ do not directly translate into observable quantities like accuracy if the Tsybakov condition is not satisfied,
and even if it is satisfied, the usefulness depends on the constants $C$ and $\lambda$.
Conversely, the following theorem makes \emph{generally} applicable statement about the accuracy that can be achieved.
Before we get to the theorem, we introduce some notation.
We are given a dataset $\mathcal{D} = {\{(x_i, y_i)\}}_i$,
where the $x_i$ are vectors of features and the $y_i$ the corresponding labels.
We refer to the tuples $(x, y)$ as the \emph{samples} of the dataset.
The number of samples is $N = |\mathcal{D}|$.

@@

The fair GP models and the baseline GP model are all based on variational inference and use the same settings.
During training, each batch is equivalent to the whole dataset.
The number of inducing inputs is 500 on the ProPublica dataset
and 2500 on the Adult dataset
which corresponds to approximately $\nicefrac{1}{8}$ of the number of training points for each dataset.
We use a squared-exponential (SE) kernel with automatic relevance determination (ARD)
and the probit function as the likelihood function.
We optimise the hyper-parameters and the variational parameters
using the Adam method~\citep{kingma2015adam} with the default parameters. %($\beta_1 = 0.9$, $\beta_2 = 0.999$).
We use the full covariance matrix for the Gaussian variational distribution.
%
The logistic regression is trained with RAdam~\citep{liu2019variance} and uses L2 regularisation.
% The regularization coefficient differs between datasets but not between repeats.
For the regularisation coefficient, we conducted a hyper-parameter search over 10 folds of the data.
For each fold, we picked the hyper-parameter which achieved the best fairness among those 5 with
the best accuracy scores.
We then averaged over the 10 hyper-parameter values chosen in this way and then used this average for all runs to obtain our final results.
% For the Adult dataset, the regularization parameter is $0.0024$ for ProPublica and $0.00035$ for Adult.

@@

Here, we have developed a machine learning framework which allows us to learn from an implicit balanced dataset,
thus satisfying the two most popular notions of fairness \citep{verma2018fairness}, demographic parity
(also known as \emph{avoiding disparate treatment}) and equality of opportunity (or \emph{avoiding disparate mistreatment}).
Additionally, we indicate how to extend the framework to cover conditional demographic parity as well.
The framework allows us to set a \emph{target rate} to control how the fairness constraint is realised.
%set a \emph{target rate} for a variety of fairness notions, including demographic parity and equality of opportunity. %equalized odds, 
For example, we can set the target positive rate for demographic parity to be $0.6$ for different groups.
Depending on the application, it can be important to specify
whether non-discrimination ought to be achieved by more positive predictions or more negative predictions.
%
This capability is unique to our approach and can be used as an intuitive mechanism to control the realisation of fairness.
Our framework is general and will be applicable for sensitive variables with binary and multi-level values.
The current work focuses on a single binary sensitive variable. %, but can be extended
Future work could extend our tuning approach to other fairness concepts like the closely related predictive parity group fairness~\citep{chouldechova2017fair}
or individual fairness~\citep{dwork2012fairness}.

>>>>

There are many existing models for enforcing demographic parity and equalised odds
\citep{creager2019flexibly,AgaBeyDudLanetal18,calders2009building,kamishima2012fairness,zafar2017fairnesstreatment,zafar2017fairnessconstraints}.
However, these existing approaches to balancing accuracy and fairness rely on intermediate, unintuitive control parameters such as allowable constraint violation $\epsilon$ (e.g. $0.01$) in \citet{AgaBeyDudLanetal18}, or a covariance threshold $c$
(e.g. $0$ that is controlled by other parameters $\tau$ and $\mu$ -- $0.005$ and $1.2$ -- to trade off this threshold and accuracy)
in \citet{zafar2017fairnesstreatment}.
This is related to the fact that many of these approaches embed fairness criteria as \emph{constraints} in the optimisation procedure
\citep{DonOneBenShaetal18,quadrianto2017recycling,zafar2017fairnesstreatment,zafar2017fairnessconstraints}.

@@

\section{Target labels for tuning group fairness}
We will start by describing several notions of group fairness.
For each individual, we have a vector of non-sensitive attributes $x\in\mathcal{X}$, a class label $y\in\mathcal{Y}$, and a sensitive attribute $s\in\mathcal{S}$ (e.g.\ racial origin or gender).
%
We focus on the case where $s$ and $y$ are binary. 
%
We assume that a positive label $y=1$ corresponds to a positive outcome for an individual -- for example, being accepted for a loan.
%
%
\emph{Group fairness} balances a certain condition between groups of individuals with different sensitive attributes, $s$ versus $s\sp{\prime}$. 
%
The term $\hat{y}$ below is the prediction of a machine learning model that,
in most works, uses only non-sensitive attributes $x$.
%
Several group fairness criteria have been proposed (e.g. \cite{zafar2017fairnesstreatment,chouldechova2017fair,hardt2016equality}):
\begin{align}
& \text{equality of positive rate (Demographic Parity):}\nonumber\\
& P(\hat{y}=1|s)=P(\hat{y}=1|s\sp{\prime}) \label{eq:eq_dem_par}\\
& \text{equality of accuracy:}\nonumber\\
& P(\hat{y}=y|s)=P(\hat{y}=y|s\sp{\prime}) \\
& \text{equality of true positive rate (Equality of Opportunity):} \nonumber\\
& P(\hat{y}=1|s,y=1)=P(\hat{y}=1|s\sp{\prime},y=1)  \label{eq:eq_opp_cri}~.
\end{align}
\emph{Equalised odds} criterion corresponds to Equality of Opportunity (\ref{eq:eq_opp_cri}) plus equality of false positive rate.

@@

%We now present a novel method for learning from an only implicitly-defined balanced dataset.
We present a novel approach which only implicitly constructs the balanced dataset.
This framework can be used with any likelihood-based model, such as Logistic Regression and Gaussian Process models.
The relation presented in \eqref{eq:etaeta} allows us to formulate a likelihood
that targets $\bar{\eta}_s(x)$ while only having access to the imbalanced labels $y$.
As we only have access to $y$, $P(y|x,s,\theta)$ is the likelihood to optimise.
It represents the probability that $y$ is the imbalanced label,
given the input $x$, the sensitive attribute $s$ that is available in the training set
and the model parameters $\theta$ for a model that is targeting $\bar{y}$.
Thus, we get
\begin{align}
  &P(y=1|x, s, \theta)
  = \sum\limits_{\bar{y}\in \{0, 1\}} P(y=1,\bar{y}|x, s, \theta) \nonumber\\
  =\,\, &\sum\limits_{\bar{y}\in \{0, 1\}} P(y=1|\bar{y}, x, s, \theta)\,P(\bar{y}|x, s, \theta) \label{eq:lik}~.
\end{align}
As we are only considering group fairness, we have $P(y=1|\bar{y}, x, s, \theta)=P(y=1|\bar{y}, s)$.

@@

The previous statements about $t_s$ do not directly translate into observable quantities like accuracy if the Tsybakov condition is not satisfied,
and even if it is satisfied, the usefulness depends on the constants $C$ and $\lambda$.
Conversely, the following theorem makes a \emph{generally} applicable statement about the accuracy that can be achieved.
Before we get to the theorem, we introduce some notation.
We are given a dataset $\mathcal{D} = {\{(x_i, y_i)\}}_i$,
where the $x_i$ are vectors of features and the $y_i$ the corresponding labels.
We refer to the tuples $(x, y)$ as the \emph{samples} of the dataset.
The number of samples is $N = |\mathcal{D}|$.

@@

The fair GP models and the baseline GP model are all based on variational inference and use the same settings.
During training, each batch is equivalent to the whole dataset.
The number of inducing inputs is 500 on the ProPublica dataset
and 2500 on the Adult dataset
which corresponds to approximately $\nicefrac{1}{8}$ of the number of training points for each dataset.
We use a squared-exponential (SE) kernel with automatic relevance determination (ARD)
and the probit function as the likelihood function.
We optimise the hyper-parameters and the variational parameters
using the Adam method~\citep{kingma2015adam} with the default parameters. %($\beta_1 = 0.9$, $\beta_2 = 0.999$).
We use the full covariance matrix for the Gaussian variational distribution.
%
The logistic regression is trained with RAdam~\citep{liu2019variance} and uses L2 regularisation.
% The regularisation coefficient differs between datasets but not between repeats.
For the regularisation coefficient, we conducted a hyper-parameter search over 10 folds of the data.
For each fold, we picked the hyper-parameter which achieved the best fairness among those 5 with
the best accuracy scores.
We then averaged over the 10 hyper-parameter values chosen in this way and then used this average for all runs to obtain our final results.
% For the Adult dataset, the regularisation parameter is $0.0024$ for ProPublica and $0.00035$ for Adult.

@@

Here, we have developed a machine learning framework which allows us to learn from an implicit balanced dataset,
thus satisfying the two most popular notions of fairness \citep{verma2018fairness}, demographic parity
(also known as \emph{avoiding disparate treatment}) and equality of opportunity (or \emph{avoiding disparate mistreatment}).
Additionally, we indicate how to extend the framework to cover conditional demographic parity as well.
The framework allows us to set a \emph{target rate} to control how the fairness constraint is realised.
%set a \emph{target rate} for a variety of fairness notions, including demographic parity and equality of opportunity. %equalised odds, 
For example, we can set the target positive rate for demographic parity to be $0.6$ for different groups.
Depending on the application, it can be important to specify
whether non-discrimination ought to be achieved by more positive predictions or more negative predictions.
%
This capability is unique to our approach and can be used as an intuitive mechanism to control the realisation of fairness.
Our framework is general and will be applicable for sensitive variables with binary and multi-level values.
The current work focuses on a single binary sensitive variable. %, but can be extended
Future work could extend our tuning approach to other fairness concepts like the closely related predictive parity group fairness~\citep{chouldechova2017fair}
or individual fairness~\citep{dwork2012fairness}.

### 2021-07-06 paper2/appendix.tex
Lots of small improvements
Fix all minor issues

\begin{table}[tbp]
    \caption{
        Additional fairness metrics for the experiments on the CelebA dataset (fig.~\ref{fig:celeba-multiplot} from the main text).
        \emph{TPR diff.} refers to the difference in true positive rate.
        \emph{TNR diff.} refers to the difference in true negative rate.
        \textsc{Left:} $\eta = 0$. \textsc{Right:} $\eta=1$.
    }
    \label{tab:my_label}
    \resizebox{.49\textwidth}{!}{
    \begin{tabular}{lrrrr}
\toprule
     Method &  Accuracy &  DP diff &  TPR diff &  TNR diff \\
\midrule
      cFlow &      0.83 &     0.10 &      0.15 &      0.25 \\
       cVAE &      0.82 &     0.05 &      0.09 &      0.18 \\
        CNN &      0.61 &     0.63 &      0.70 &      0.64 \\
 Ln2L &      0.52 &     0.00 &      0.00 &      0.00 \\
\bottomrule
\end{tabular}}
\hfill
\resizebox{.49\textwidth}{!}{
\begin{tabular}{lrrrr}
\toprule
     Method &  Accuracy &  DP diff &  TPR diff &  TNR diff \\
\midrule
      cFlow &      0.82 &     0.33 &      0.28 &      0.21 \\
       cVAE &      0.81 &     0.16 &      0.10 &      0.05 \\
        CNN &      0.67 &     0.75 &      0.66 &      0.76 \\
Ln2L &      0.51 &     0.08 &      0.06 &      0.09 \\
\bottomrule
\end{tabular}}
\end{table}
\paragraph{Additional fairness metrics.}
In addition to \emph{DP diff}, we report here the result from other fairness measures.
These results are from the same setup as those reported in the main paper.
We report the difference in true positive rates (TPR) between the two groups (male and female), which corresponds to a measure of Equality of Opportunity,
and the difference in true negative rates (TNR) between the two groups.

@@

\paragraph{Tuning the Partition Sizes.}
There are several ways of ensuring that $z_b$ is of sufficient size to capture all $s$ dependencies, while at time is minimised such as that information unrelated to $s$ is maximally preserved. We adopt the straightforward search strategy of, starting from some initial guess, calibrating the value according to accuracy attained by a classifier trained to predict $s$ from $z_b$ on a held-out subset of the representative set, which is measured whenever the adversarial loss plateaus. If the accuracy is above chance level then that suggests the size of the $z_b$ partition, $|z_b|$, needs to be increased to accommodate more information about $s$. If the accuracy is found to be at chance level then are two possibilities: 1) $|z_b|$ is already optimal; 2) $|z_b|$ is large enough that it fully contains both information $s$ as well as that of a portion of $y$. If the former is true, then perturbations around the current value allow us to confirm this; if the latter is true then decreasing the value was indeed the correct decision.

>>>>

\begin{table}[tbp]
    \caption{
        Additional fairness metrics for the experiments on the CelebA dataset (fig.~\ref{fig:celeba-multiplot} from the main text).
        \emph{TPR diff.}\ refers to the difference in true positive rate.
        \emph{TNR diff.}\ refers to the difference in true negative rate.
        \textsc{Left:} $\eta = 0$. \textsc{Right:} $\eta=1$.
    }
    \label{tab:my_label}
    \resizebox{.49\textwidth}{!}{
    \begin{tabular}{lrrrr}
\toprule
     Method &  Accuracy &  DP diff &  TPR diff &  TNR diff \\
\midrule
      cFlow &      0.83 &     0.10 &      0.15 &      0.25 \\
       cVAE &      0.82 &     0.05 &      0.09 &      0.18 \\
        CNN &      0.61 &     0.63 &      0.70 &      0.64 \\
 Ln2L &      0.52 &     0.00 &      0.00 &      0.00 \\
\bottomrule
\end{tabular}}
\hfill
\resizebox{.49\textwidth}{!}{
\begin{tabular}{lrrrr}
\toprule
     Method &  Accuracy &  DP diff &  TPR diff &  TNR diff \\
\midrule
      cFlow &      0.82 &     0.33 &      0.28 &      0.21 \\
       cVAE &      0.81 &     0.16 &      0.10 &      0.05 \\
        CNN &      0.67 &     0.75 &      0.66 &      0.76 \\
Ln2L &      0.51 &     0.08 &      0.06 &      0.09 \\
\bottomrule
\end{tabular}}
\end{table}
\paragraph{Additional fairness metrics.}
In addition to \emph{DP diff}, we report here the result from other fairness measures.
These results are from the same setup as those reported in the main paper.
We report the difference in true positive rates (TPR) between the two groups (male and female), which corresponds to a measure of Equality of Opportunity,
and the difference in true negative rates (TNR) between the two groups.

@@

\paragraph{Tuning the Partition Sizes.}
There are several ways of ensuring that the size of $z_b$ is sufficient to capture all s dependencies, but minimal enough that information unrelated to s is maximally preserved
We adopt the straightforward search strategy of, starting from some initial guess, calibrating the value according to accuracy attained by a classifier trained to predict $s$ from $z_b$ on a held-out subset of the representative set, which is measured whenever the adversarial loss plateaus. If the accuracy is above chance level then that suggests the size of the $z_b$ partition, $|z_b|$, needs to be increased to accommodate more information about $s$. If the accuracy is found to be at chance level then are two possibilities: 1) $|z_b|$ is already optimal; 2) $|z_b|$ is large enough that it fully contains both information $s$ as well as that of a portion of $y$. If the former is true, then perturbations around the current value allow us to confirm this; if the latter is true then decreasing the value was indeed the correct decision.

### 2021-07-06 paper3/paper3.tex
Fix all minor issues

\section{Abstract}
Machine learning models are typically trained to optimise global metrics such as average classification accuracy. 
%
Hidden stratification arises when the trained models have high average performance over classes but exhibit highly variable performance across different hidden subgroups.
%
In this paper, we consider the setting where the hidden stratification has zero class-labelled data for some subgroups. 
As an illustration, we have digit images labelled as "two" or "four", each class comprises "green" and "purple" subgroups.
%
Challengingly, in the training data, twos can be any colour, but all fours are green. 
%
Without additional knowledge, it is impossible to directly control the discrepancy of the classifier’s statistics for the hidden subgroups. 
%
We develop a disentanglement algorithm that decomposes a data representation into a component that captures the subgrouping factors and a component that is invariant to them based on unlabelled (deployment) data. 
%
We cluster the unlabelled data, and equalise the cluster sizes to form "perfect bags" with respect to class and subgroup information.
%
We cast the problem of disentangling as one of distribution matching and propose an adversarial learning approach.
%
%We introduce a new discriminator loss, inspired by set-classification, to distinguish a batch of perfect bags from non-perfect ones based on a learnable attention mechanism. 
%
Unlike sample-based models, we advance a discriminator to assign scores at the level of bags of samples, with a bag being deemed authentic if was drawn from the unbiased distribution. 
%a parametric approach based on adversarial learning. 
%We take inspiration from set-classification and equip the discriminator with a learnable attention mechanism to model interdependencies between samples in a bag.
%
We evaluate our approach on several classification benchmarks and show that it is indeed possible to account for zero-label hidden stratification.
% \end{abstract}

\section{Introduction}%
\label{sec:introduction}
Machine learning has been deployed in safety-critical applications such as medicine (e.g. \citep{DunYiLanReetal19}), and socially important contexts such as the allocation of healthcare, education, and credit (\eg\ \citep{HurAde17,RagBarKleLev20}). 
%
Efficiency can be improved, costs can be reduced, and personalisation of services and products can be greatly enhanced -- these are some of the drivers for the widespread development and deployment of machine learning algorithms. 
%

Algorithms such as classifiers, however, are trained from large amount of labelled data, and are typically trained to optimise \emph{global} metrics such as average classification accuracy.
%
% \begin{figure*}[t]
% \centering
% \includegraphics[width=0.6\textwidth,page=2]{paper3/figures/ideal.pdf}
% \caption{Overview of the zero-shot stratification problem. Training dataset (pairs of input data $x$ and class label $y$) will only contain data points that are labelled ``yes'' by the decision policy. This systematic bias might result in a subgroup(s) to have zero labelled data (in the example above, ``married'' subgroup has zero labelled data). The subgrouping information $s$ is unavailable/unlabelled at deployment time, and is only partially labelled at train time.}
% \vspace{-0.5cm}
% \label{fig:censoring}
% \end{figure*}
In many real-world classification tasks, each labelled class consists of multiple semantically distinct subclasses, or subgroups.
For example, the ``dog'' class label can have finer-grained intra-class variations, such as ``dog indoor'' and ``dog outdoor''.
%
This finer-grained subgrouping information is typically unavailable/unlabelled (e.g. \citep{SohDunAngGuetal20,nam2020learning}).
%
The standard training process brings about two inter-connected challenges: a) classifiers often under-perform on important hidden subgroups (\emph{hidden stratification}) \citep{RayDunCarRe20,SohDunAngGuetal20};
and b) \emph{systematic bias}~\citep{kallus2018residual} affects whether or not entire collections of data points appear in the training dataset,
and can make the classifier unprepared for treating those subgroups in the eventual deployment setting
% and can hamper attempts to correct for bias in the eventual deployment data
(\emph{residual bias}). %\cite{kallus2018residual}.

@@

\paragraph{Perfect bag.}
We call a sampled set for which $y\perp s$ holds, a perfect bag \citep{chouldechova2017fair,kleinberg2016inherent}.
Such sets are also very desirable when training algorithmically fair classifiers
(see the \emph{sampling} method in \citet{kamiran2012data}).
%
%If we have access to perfect bags, we could equalise true positive/negative rates (\eqref{eq:TPR}) and also equalise positive/negative predicted values (\eqref{eq:PPV}) for all demographic groups.
%This can be shown by using the sum and product rule of conditional probabilities, e.g. \citet{KanRotZia19}.
%We consider a binary-valued subgroup label, $s\sp{\prime}$ versus $s\sp{\prime\prime}$. 
%For $s\sp{\prime}$, we can compute: 
%$P(y=1|\hat{y}=1,s\sp{\prime})=$
%%\begin{align}
%$\nicefrac{P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime})}{\left(A + B\right)}$, with $A = P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime})$ and $B = P(\hat{y}=1|y=0,s\sp{\prime})(1-P(y=1|s\sp{\prime}))$.
%%\end{align}
%We can do similarly for $s\sp{\prime\prime}$.
%The conditional probability on the left hand side is 
%a positive predicted value, and this quantity can be expressed in terms of true positive/negative rates and the base (prior) rate, shown on the right hand side. 
%If we have a perfect bag ($y\perp s$ holds, which means equal base rates $P(y=1|s\sp{\prime}) = P(y=1|s\sp{\prime\prime})$), an equality in the true positive/negative rates will give us an equality in the positive/negative predicted values.
%Similarly, with a perfect bag, we can equalise true positive/negative rates (eq. \ref{eq:TPR}) and also positive rates (eq. \ref{eq:AR}) for all demographic groups.
%From the sum probability rule, we have:
%$
%P(\hat{y}=1|s\sp{\prime})
%= P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime}) + P(\hat{y}=1|y=0,s\sp{\prime})(1-P(y=1|s\sp{\prime}))
%$ for $s\sp{\prime}$ value, and accordingly for $s\sp{\prime\prime}$ value.
%Here, a positive rate on the left hand side is related to true positive/negative rates and the base (prior) rate as shown on the right hand side. 
%
Ideally, we would like to sample our deployment dataset as perfect bags.
However, the deployment set is unlabelled and is unlikely to be perfect in practice. 
Instead, we pursue learning under zero-label systematic bias as learning disentangled representations with a collection of \emph{approximately} perfect bags (produced with clustering techniques, see section~\ref{sec:implementation}).
We show that the disentangling procedure is robust enough to work with this relaxation, but that performance scales with how well the deployment set is balanced.

@@

\paragraph{Clustering guarantees.}
By leveraging the feature space of deep models, previous research has shown that semantically meaningful clusters can be discovered without the need for ground-truth annotations
 (for example \citet{RayDunCarRe20,GanVanGeoProetal20,SohDunAngGuetal20,HanRebEhrVedetal20}). 
\citet{SohDunAngGuetal20} recently provided a ``clustering guarantee'' when assuming access to ground truth class labels $y$ and using clustering within each class to generate approximate subgroups $s$.  
%
Even with infinite data, accurate identification of the subgroups is unfortunately impossible. 
%
Instead, \citet{SohDunAngGuetal20} showed that the error in their quantity of interest, classification loss for each subgroup, can be bounded by the total variation estimation error in the mixture-of-Gaussians case.   
%
Our primary goal is to ensure good construction of perfect bags.
We can adapt the analysis of \citet{SohDunAngGuetal20} to bound the error of our quantity of interest, \emph{aggregate statistics} of the approximated perfect bag for each subgroup.
%
We can subsequently apply the union bound to take into account that we have a fully-unlabelled setting.

>>>>

\section{Abstract}
Machine learning models are typically trained to optimise global metrics such as average classification accuracy. 
%
Hidden stratification arises when the trained models have high average performance over classes but exhibit highly variable performance across different hidden subgroups.
%
In this paper, we consider the setting where the hidden stratification has zero class-labelled data for some subgroups. 
As an illustration, we have digit images labelled as ``two'' or ``four'', each class comprises ``green'' and ``purple'' subgroups.
%
Challengingly, in the training data, twos can be any colour, but all fours are green. 
%
Without additional knowledge, it is impossible to directly control the discrepancy of the classifier’s statistics for the hidden subgroups. 
%
We develop a disentanglement algorithm that decomposes a data representation into a component that captures the subgrouping factors and a component that is invariant to them based on unlabelled (deployment) data. 
%
We cluster the unlabelled data, and equalise the cluster sizes to form ``perfect bags'' with respect to class and subgroup information.
%
We cast the problem of disentangling as one of distribution matching and propose an adversarial learning approach.
%
%We introduce a new discriminator loss, inspired by set-classification, to distinguish a batch of perfect bags from non-perfect ones based on a learnable attention mechanism. 
%
Unlike sample-based models, we advance a discriminator to assign scores at the level of bags of samples, with a bag being deemed authentic if was drawn from the unbiased distribution. 
%a parametric approach based on adversarial learning. 
%We take inspiration from set-classification and equip the discriminator with a learnable attention mechanism to model interdependencies between samples in a bag.
%
We evaluate our approach on several classification benchmarks and show that it is indeed possible to account for zero-label hidden stratification.
% \end{abstract}

\section{Introduction}%
\label{sec:introduction}
Machine learning has been deployed in safety-critical applications such as medicine \citep[e.g.][]{DunYiLanReetal19}, and socially important contexts such as the allocation of healthcare, education, and credit \citep[e.g.][]{HurAde17,RagBarKleLev20}.
%
Efficiency can be improved, costs can be reduced, and personalisation of services and products can be greatly enhanced -- these are some of the drivers for the widespread development and deployment of machine learning algorithms. 
%

Algorithms such as classifiers, however, are trained from large amount of labelled data, and are typically trained to optimise \emph{global} metrics such as average classification accuracy.
%
% \begin{figure*}[t]
% \centering
% \includegraphics[width=0.6\textwidth,page=2]{paper3/figures/ideal.pdf}
% \caption{Overview of the zero-shot stratification problem. Training dataset (pairs of input data $x$ and class label $y$) will only contain data points that are labelled ``yes'' by the decision policy. This systematic bias might result in a subgroup(s) to have zero labelled data (in the example above, ``married'' subgroup has zero labelled data). The subgrouping information $s$ is unavailable/unlabelled at deployment time, and is only partially labelled at train time.}
% \vspace{-0.5cm}
% \label{fig:censoring}
% \end{figure*}
In many real-world classification tasks, each labelled class consists of multiple semantically distinct subclasses, or subgroups.
For example, the ``dog'' class label can have finer-grained intra-class variations, such as ``dog indoor'' and ``dog outdoor''.
%
This finer-grained subgrouping information is typically unavailable/unlabelled \citep[e.g.][]{SohDunAngGuetal20,nam2020learning}.
%
The standard training process brings about two inter-connected challenges: a) classifiers often under-perform on important hidden subgroups (\emph{hidden stratification}) \citep{RayDunCarRe20,SohDunAngGuetal20};
and b) \emph{systematic bias}~\citep{kallus2018residual} affects whether or not entire collections of data points appear in the training dataset,
and can make the classifier unprepared for treating those subgroups in the eventual deployment setting
% and can hamper attempts to correct for bias in the eventual deployment data
(\emph{residual bias}). %\cite{kallus2018residual}.

@@

\paragraph{Perfect bag.}
We call a sampled set for which $y\perp s$ holds, a perfect bag \citep{chouldechova2017fair,kleinberg2016inherent}.
Such sets are also very desirable when training algorithmically fair classifiers
\citep[see the \emph{sampling} method in][]{kamiran2012data}.
%
%If we have access to perfect bags, we could equalise true positive/negative rates (\eqref{eq:TPR}) and also equalise positive/negative predicted values (\eqref{eq:PPV}) for all demographic groups.
%This can be shown by using the sum and product rule of conditional probabilities, e.g. \citet{KanRotZia19}.
%We consider a binary-valued subgroup label, $s\sp{\prime}$ versus $s\sp{\prime\prime}$. 
%For $s\sp{\prime}$, we can compute: 
%$P(y=1|\hat{y}=1,s\sp{\prime})=$
%%\begin{align}
%$\nicefrac{P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime})}{\left(A + B\right)}$, with $A = P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime})$ and $B = P(\hat{y}=1|y=0,s\sp{\prime})(1-P(y=1|s\sp{\prime}))$.
%%\end{align}
%We can do similarly for $s\sp{\prime\prime}$.
%The conditional probability on the left hand side is 
%a positive predicted value, and this quantity can be expressed in terms of true positive/negative rates and the base (prior) rate, shown on the right hand side. 
%If we have a perfect bag ($y\perp s$ holds, which means equal base rates $P(y=1|s\sp{\prime}) = P(y=1|s\sp{\prime\prime})$), an equality in the true positive/negative rates will give us an equality in the positive/negative predicted values.
%Similarly, with a perfect bag, we can equalise true positive/negative rates (eq. \ref{eq:TPR}) and also positive rates (eq. \ref{eq:AR}) for all demographic groups.
%From the sum probability rule, we have:
%$
%P(\hat{y}=1|s\sp{\prime})
%= P(\hat{y}=1|y=1,s\sp{\prime})P(y=1|s\sp{\prime}) + P(\hat{y}=1|y=0,s\sp{\prime})(1-P(y=1|s\sp{\prime}))
%$ for $s\sp{\prime}$ value, and accordingly for $s\sp{\prime\prime}$ value.
%Here, a positive rate on the left hand side is related to true positive/negative rates and the base (prior) rate as shown on the right hand side. 
%
Ideally, we would like to sample our deployment dataset as perfect bags.
However, the deployment set is unlabelled and is unlikely to be perfect in practice. 
Instead, we pursue learning under zero-label systematic bias as learning disentangled representations with a collection of \emph{approximately} perfect bags (produced with clustering techniques, see section~\ref{sec:implementation}).
We show that the disentangling procedure is robust enough to work with this relaxation, but that performance scales with how well the deployment set is balanced.

@@

\paragraph{Clustering guarantees.}
By leveraging the feature space of deep models, previous research has shown that semantically meaningful clusters can be discovered without the need for ground-truth annotations
\citep[for example][]{RayDunCarRe20,GanVanGeoProetal20,SohDunAngGuetal20,HanRebEhrVedetal20}.
\citet{SohDunAngGuetal20} recently provided a ``clustering guarantee'' when assuming access to ground truth class labels $y$ and using clustering within each class to generate approximate subgroups $s$.  
%
Even with infinite data, accurate identification of the subgroups is unfortunately impossible. 
%
Instead, \citet{SohDunAngGuetal20} showed that the error in their quantity of interest, classification loss for each subgroup, can be bounded by the total variation estimation error in the mixture-of-Gaussians case.   
%
Our primary goal is to ensure good construction of perfect bags.
We can adapt the analysis of \citet{SohDunAngGuetal20} to bound the error of our quantity of interest, \emph{aggregate statistics} of the approximated perfect bag for each subgroup.
%
We can subsequently apply the union bound to take into account that we have a fully-unlabelled setting.
